% CREATED BY MAGNUS GUSTAVER, 2020
\chapter{Experiment and Analysis}
\label{chap:experiment-analysis}
\label{chap:results}

This chapter evaluates NanoMem as a memory evidence provider. The central
question is whether the system can construct compact, source-grounded, and
temporally grounded evidence for long-term conversational memory questions. The
chapter first describes the experimental setup, including benchmarks,
baselines, metrics, and training/evaluation boundaries. It then reports the main
benchmark results on LoCoMo, LongMemEval, and TimeMemEval, followed by a
comparative analysis across baseline families. The final sections analyze
architecture and reward ablations to explain which parts of NanoMem account for
the observed behavior.

\section{Experimental Setup}
\label{sec:experiment-setup}

\subsection{Benchmarks}
\label{subsec:experiment-benchmarks}

We evaluate NanoMem on two public long-term memory benchmarks and one diagnostic
benchmark constructed for this thesis. The three benchmarks are complementary:
LoCoMo and LongMemEval establish whether NanoMem remains effective on standard
long-term conversational memory tasks, while TimeMemEval isolates the
hidden-dependency temporal bridge setting that motivates the method.

\textbf{LoCoMo} evaluates long-term multi-session dialogue memory with
single-hop, multi-hop, temporal reasoning, and open-domain questions
\cite{maharana2024locomo}. It is useful for testing cross-session evidence
localization and whether a memory method remains accurate when relevant facts
are distributed across long dialogue histories.

\textbf{LongMemEval} focuses on long-term memory for chat assistants and
contains single-session user/assistant questions, knowledge updates, temporal
reasoning, and multi-session reasoning \cite{wu2024longmemeval}. Compared with
LoCoMo, its histories are noisier and closer to assistant-style long-memory
usage, making it suitable for testing whether compact evidence construction can
outperform direct full-context prompting.

\textbf{TimeMemEval} is the diagnostic benchmark introduced in this thesis. Each
instance requires a model to find a source event, derive a target temporal
window, retrieve a destination event in that window, and answer a terminal
question about the destination event. The benchmark is not intended to cover all
temporal reasoning phenomena. Instead, it specifically tests the evidence
addressability gap: the original query does not directly contain the retrieval
cue needed to find the final answer-bearing evidence.

The training and evaluation split is reported in
Appendix~\ref{app:training-evaluation-split}. In brief, GRPO training uses 500
examples from LoCoMo and TimeMemEval, while LongMemEval is reserved for
zero-shot evaluation.

\subsection{Baselines}
\label{subsec:experiment-baselines}

We compare NanoMem with four groups of baselines. Full-context baselines feed
the complete available history directly to GPT-5.1 or Qwen3.5-9B. These rows
serve as strong long-context references, but they are not memory systems and do
not represent the lowest-cost deployment option. Write-time memory systems,
including Mem0 \cite{chhikara2025mem0} and A-Mem \cite{xu2025amem}, construct
or organize memory records before the future query is known. Search-time
retrieval agents, including MemR$^3$ \cite{du2025memr3} and Search-R1
\cite{jin2025searchr1}, retrieve or search at query time but do not maintain
NanoMem's structured temporal evidence pool. The temporal baseline is
Memory-T1-Qwen \cite{du2026memoryt1}. Since the official trained Memory-T1
checkpoint was not available, this row is a Memory-T1-style implementation
using Qwen3.5-9B rather than official Memory-T1 checkpoint performance.

NanoMem is reported in two variants. NanoMem-Vanilla uses the same
Planner--Retriever--Synthesizer framework without GRPO training. NanoMem-GRPO
uses the trained Synthesizer. The comparison between the two variants separates
the contribution of the framework from the learned evidence construction
policy.

\subsection{Metrics and Evaluation Protocol}
\label{subsec:experiment-metrics}

The main metric is answer accuracy. A fixed GPT-5.1 judge compares each
predicted answer with the gold answer under the same judge prompt across all
methods. We do not use BLEU or F1 as the primary metric because long-term memory
answers often admit semantically equivalent natural-language variants,
especially for temporal expressions and event descriptions.

Token counts are reported as output-side compactness metrics and must be read
carefully. For direct-answerer methods, tokens denote answer-output tokens. For
evidence-provider methods such as NanoMem, Mem0, A-Mem, and MemR$^3$, tokens
denote retrieved evidence or memory context returned downstream. They are
therefore useful for comparing evidence compactness, but they are not identical
to full input cost for every method.

For NanoMem-specific diagnostics, we additionally report retrieval rounds and
verdict accuracy. Retrieval rounds measure how often the iterative loop is used.
Fewer rounds are not automatically better: the desired behavior is to stop when
the evidence is sufficient and continue when a missing bridge must be recovered.
Verdict accuracy measures whether the Synthesizer's sufficient/insufficient
decision agrees with gold evidence coverage.

During training, the Qwen3.5-9B Synthesizer is optimized with full-parameter
GRPO using VERL/HybridFlow \cite{sheng2024hybridflow}. The Planner, NanoMem
retriever, Qwen3.5-9B downstream answerer, GPT-5.1 judge, and reference policy
are frozen environment components. Each prompt samples \(K=8\) trajectories with
at most three retrieval rounds. At inference time, NanoMem uses the same
three-round limit and retrieves the top-7 memories in each round. The final run
uses reward weights \((w_o,w_v,w_l)=(0.50,0.10,0.05)\) and a format-failure
penalty \(c=-0.20\). Training is conducted on four NVIDIA GH200 96GB GPUs.

\section{Benchmark Results}
\label{sec:benchmark-results}
\label{app:main-benchmark-results}

Table~\ref{tab:main-summary} summarizes the main accuracy results
across LoCoMo, LongMemEval, and TimeMemEval. NanoMem-GRPO obtains the best
overall accuracy on all three benchmarks: 88.6\% on LoCoMo, 83.8\% on
LongMemEval, and 51.5\% on TimeMemEval. Compared with NanoMem-Vanilla, GRPO
raises the overall score from 85.0\% to 88.6\% on LoCoMo, from 78.7\% to
83.8\% on LongMemEval, and from 43.0\% to 51.5\% on TimeMemEval. The largest
relative gain appears on TimeMemEval, where evidence is not directly
addressable from the original query and the system must recover an intermediate
temporal bridge before continuing search.

\input{tbl/main_summary}

\subsection{LoCoMo}
\label{subsec:results-locomo}

On LoCoMo, NanoMem-GRPO reaches 88.6\% overall accuracy, slightly above the
GPT-5.1 full-context reference at 87.7\%. The category-level pattern is more
informative than the overall score alone. NanoMem-GRPO improves over
NanoMem-Vanilla on multi-hop questions from 84.0\% to 89.0\%, and on temporal
questions from 76.6\% to 87.5\%. It also remains strong on single-hop questions
at 91.4\%, showing that the temporal and multi-hop optimization does not break
ordinary memory recall. The open-domain score is 65.2\%, slightly below
NanoMem-Vanilla's 67.4\%, so the correct interpretation is not that every
category improves uniformly. The gain is concentrated in the multi-hop and
temporal categories targeted by the method.

\input{tbl/locomo_main}

The token counts also show that the improvement is not caused by returning a
larger context. NanoMem-GRPO returns 107 tokens on average, compared with 264
for NanoMem-Vanilla, 650 for Mem0, and 12.3k for A-Mem. On LoCoMo, the trained
Synthesizer therefore improves accuracy while producing substantially more
compact evidence.

\subsection{LongMemEval}
\label{subsec:results-longmemeval}

The gap between full-context prompting and evidence construction becomes larger
on LongMemEval. NanoMem-GRPO reaches 83.8\% overall accuracy, compared with
63.6\% for GPT-5.1 full context. This should not be read as a general claim
that GPT-5.1 is weak. Instead, it shows that in long and noisy histories,
directly asking the answer model to process the whole context can be less
effective than selecting and synthesizing the relevant evidence first.

The temporal reasoning category is the clearest result. NanoMem-GRPO reaches
86.4\% on LongMemEval-TR, compared with 54.2\% for Mem0, 55.0\% for
MemR$^3$, and 27.1\% for Memory-T1-Qwen. NanoMem-GRPO also remains strong on
single-session categories, reaching 95.7\% on SSU and 98.2\% on SSA. The
multi-session reasoning score is 71.4\%, slightly above NanoMem-Vanilla at
70.7\%.

\input{tbl/longmemeval_main}

LongMemEval also confirms the compactness pattern. NanoMem-GRPO returns 119
tokens on average, compared with 353 for NanoMem-Vanilla, 217 for Mem0, and
18.1k for A-Mem. Thus, the improvement over full-context prompting and
write-time memory systems is paired with shorter evidence rather than larger
returned context.

\subsection{TimeMemEval}
\label{subsec:results-timememeval}

TimeMemEval is harder in absolute terms, which is expected because it is a
diagnostic temporal-bridge benchmark. NanoMem-GRPO reaches 51.5\% overall
accuracy, compared with 43.0\% for NanoMem-Vanilla. The normal split reaches
57.0\% and the hard split reaches 47.0\%. Full-context baselines remain much
lower: GPT-5.1 full context reaches 16.0\%, and Qwen3.5-9B full context reaches
20.5\%. Existing memory and retrieval baselines are also far below NanoMem on
this diagnostic task, with Mem0 at 1.5\%, A-Mem at 7.5\%, MemR$^3$ at 12.0\%,
Search-R1 at 6.5\%, and Memory-T1-Qwen at 3.5\%.

\input{tbl/chainmem_main}

The absolute accuracy remains far from saturated, which is an important part of
the result. TimeMemEval is intentionally difficult: the final evidence is hidden
behind an intermediate event, time window, or condition. NanoMem does not solve
this problem completely, but it improves substantially over both full-context
and existing memory baselines. This supports the central claim that temporal
bridge search requires more than one-shot semantic matching.

\section{Comparative Analysis}
\label{sec:comparative-analysis}

\subsection{Full-context Prompting versus Memory Evidence Construction}
\label{subsec:analysis-full-context}

Full-context prompting is a useful reference because it gives the answer model
access to the full available history. However, the results show that access is
not the same as usable evidence. On LoCoMo, NanoMem-GRPO is close to the
GPT-5.1 full-context reference and slightly exceeds it overall, 88.6\% versus
87.7\%, while returning compact evidence rather than relying on the complete
dialogue history. On LongMemEval, the gap is larger: NanoMem-GRPO reaches
83.8\%, while GPT-5.1 full context reaches 63.6\%. On TimeMemEval, GPT-5.1 full
context reaches only 16.0\%, indicating that even when the relevant sessions are
available in the prompt, the model can fail to recover the temporal bridge that
connects source and destination evidence.

The implication is not that long-context models are unhelpful. Rather,
long-term conversational memory requires evidence construction. A memory system
should identify what the downstream answerer needs to see, preserve source
grounding, and reduce irrelevant context. This is the role NanoMem is designed
to play.

\subsection{Write-time Memory versus Query-time Evidence Synthesis}
\label{subsec:analysis-write-time}

Write-time memory systems such as Mem0 and A-Mem construct memory records before
the future query is known. This is appropriate for stable user facts,
preferences, and profile-style memories. The limitation appears when the
relevant relation becomes visible only after the query is asked. TimeMemEval
exposes this limitation directly: Mem0 reaches 1.5\% and A-Mem reaches 7.5\%,
while NanoMem-GRPO reaches 51.5\%.

The token profile also shows a practical trade-off. A-Mem returns transcript-
level evidence in these experiments, with 12.3k tokens on LoCoMo, 18.1k on
LongMemEval, and 3.8k on TimeMemEval. This may preserve recall, but it shifts a
large burden to the downstream answerer. NanoMem instead performs lightweight
write-time processing and uses query-time evidence synthesis to decide what is
relevant for the current question.

\subsection{Search-time Retrieval versus Structured Temporal Evidence State}
\label{subsec:analysis-search-time}

Search-time retrieval agents are closer to NanoMem because they react to the
current query. MemR$^3$ and Search-R1 can retrieve or search at query time, but
they do not maintain NanoMem's structured Temporal Evidence Pool or the binary
sufficient/insufficient verdict loop. This difference is most visible on
LongMemEval temporal reasoning and TimeMemEval. MemR$^3$ reaches 55.0\% on
LongMemEval-TR and 12.0\% on TimeMemEval, while NanoMem-GRPO reaches 86.4\%
and 51.5\%, respectively.

Search-R1 should be interpreted carefully. Its public checkpoint was trained for
open-domain search QA, not timestamped long-term conversational memory. The
reported row therefore evaluates a search-style baseline under a memory adapter,
not the upper bound of the Search-R1 method family. Still, the comparison
illustrates a key distinction: NanoMem's search is not only about issuing more
queries. It converts retrieved sessions into a temporally grounded evidence
state that can guide the next retrieval step.

\subsection{Accuracy and Compactness}
\label{subsec:analysis-compactness}

Compactness is not a cosmetic metric in this thesis. NanoMem is designed as an
evidence provider, so the downstream answer model should see a concise evidence
state rather than a large bag of raw sessions. Table~\ref{tab:vanilla-grpo-compactness}
compares NanoMem-Vanilla with NanoMem-GRPO. GRPO improves accuracy while
reducing returned evidence length on all three benchmarks. On LoCoMo, output
length drops from 264 to 107 tokens; on LongMemEval, from 353 to 119 tokens; on
TimeMemEval, from 287 to 234 tokens.

\begin{table}[H]
  \centering
  \caption{NanoMem-Vanilla versus NanoMem-GRPO accuracy and returned evidence
  compactness.}
  \label{tab:vanilla-grpo-compactness}
  \scriptsize
  \resizebox{\textwidth}{!}{%
  \begin{tabular}{p{0.20\textwidth}rrrrrr}
    \hline\hline
    Benchmark & Vanilla acc. & Vanilla tok. & GRPO acc. & GRPO tok. & Acc. gain & Token reduction \\
    \hline
    LoCoMo & 85.0 & 264 & 88.6 & 107 & +3.6 & 59.5\% \\
    LongMemEval & 78.7 & 353 & 83.8 & 119 & +5.1 & 66.3\% \\
    TimeMemEval & 43.0 & 287 & 51.5 & 234 & +8.5 & 18.5\% \\
    \hline\hline
  \end{tabular}
  }
\end{table}

This result should be interpreted as returned evidence compactness, not total
system cost. The input cost of each method depends on its retrieval procedure,
history length, and implementation. Nevertheless, the pattern is important:
NanoMem-GRPO improves the answer outcome without passing more evidence to the
answerer.

\section{Architecture Ablation}
\label{sec:architecture-ablation-results}
\label{sec:results-architecture}

The architecture ablation isolates three design choices: temporal text
normalization, the Temporal Evidence Pool, and multi-round retrieval. The full
row uses NanoMem-Vanilla rather than NanoMem-GRPO so that the ablation measures
framework components without mixing in the effect of the trained policy.

\input{tbl/arch_ablation}

Text normalization mainly supports temporal grounding. Removing it barely
affects LoCoMo-MH, which drops from 84.0\% to 82.3\%, but it causes larger
drops on LoCoMo-Temp, LongMemEval-TR, and TimeMemEval: 5.3, 10.9, and 9.5
points, respectively. This indicates that explicit normalization is most useful
when the query or memory contains relative time expressions, mismatched
session/event times, or a temporal bridge.

The Temporal Evidence Pool is more broadly important. Removing it lowers every
reported task score. The largest drops appear on LongMemEval-TR and
TimeMemEval, by 15.2 and 18.0 points. It also increases average rounds from
1.25 to 1.45. This means the loop continues searching, but without a structured
state of temporally resolved events, later rounds become less effective.

Single-round retrieval shows why the value of iteration is task-dependent. It
hurts LoCoMo-MH and TimeMemEval, where first-round retrieval is often
incomplete, but it slightly improves LoCoMo-Temp and LongMemEval-TR. Many
temporal questions in those categories are answerable from first-round evidence,
and extra search can add noise. NanoMem's goal is therefore not to increase the
number of rounds, but to learn when evidence is sufficient and when another
search step is necessary.

\section{Reward Ablation}
\label{sec:reward-ablation-results}
\label{sec:results-reward}

The reward ablation evaluates whether GRPO changes the Synthesizer's evidence
construction behavior. The full reward combines answer outcome, verdict
accuracy, compactness, and the format gate described in
Chapter~\ref{chap:methods}. Two variants remove different parts of this signal:
one removes the format and length rewards, and the other removes the verdict
reward.

\input{tbl/reward_ablation}

The full reward reaches 89.0\% on LoCoMo-MH, 87.5\% on LoCoMo-Temp, 86.4\% on
LongMemEval-TR, and 51.5\% on TimeMemEval. It also keeps verdict accuracy high
at 84.0\%, with only 1.10 average rounds and 110 average output tokens.

Removing the format and length rewards produces a compactness failure. Accuracy
drops moderately and consistently, but the more visible effect is output length:
average output increases from 110 to 359 tokens, a 3.3 times increase. These
rewards are therefore not merely cosmetic formatting constraints. They
regularize the Synthesizer against satisfying the outcome objective by passing
large, redundant evidence to the downstream answerer.

Removing the verdict reward exposes a different failure mode. The outcome-only
variant improves on LoCoMo-Temp and LongMemEval-TR, but it drops sharply on
LoCoMo-MH and TimeMemEval, from 89.0\% to 80.5\% and from 51.5\% to 41.5\%,
respectively. Verdict accuracy also falls from 84.0\% to 72.9\%, average rounds
increase from 1.10 to 1.25, and average tokens increase to 361. This pattern is
consistent with the task structure: when first-round evidence is often
sufficient, outcome-only training can be strong; when a multi-hop or temporal
bridge is missing, the policy needs direct supervision for the
sufficient/insufficient decision.

\section{Summary of Findings}
\label{sec:experiment-summary}

The experiments support four findings. First, NanoMem-GRPO obtains the best
overall accuracy on LoCoMo, LongMemEval, and TimeMemEval, with gains
concentrated in multi-hop, temporal reasoning, and hidden-dependency temporal
bridge settings. Second, the comparative analysis shows that full-context
prompting, write-time memory, and generic search-time retrieval do not fully
address temporal evidence addressability. NanoMem's advantage comes from
query-time evidence synthesis, the Temporal Evidence Pool, and the structured
verdict loop. Third, the compactness analysis shows that GRPO improves accuracy
without context bloat: NanoMem-GRPO returns substantially shorter evidence than
NanoMem-Vanilla on all three benchmarks. Fourth, the ablations show that the
main components play distinct roles: text normalization supports temporal
grounding, the Temporal Evidence Pool supports cross-round state transfer,
multi-round retrieval supports hidden-dependency search, and the reward
components shape compactness and sufficiency judgment.

The next chapter uses these findings to answer the research questions and
discuss limitations, including automatic judge dependence, baseline adaptation
boundaries, non-equivalent token metrics, the diagnostic scope of TimeMemEval,
and deployment concerns for long-term personal memory systems.
