% CREATED BY MAGNUS GUSTAVER, 2020
\thispagestyle{plain}			% Supress header
\section*{Acknowledgements}

I would first like to express my sincere gratitude to Mats Granath and Per
Thoren for their continuous support throughout my master's thesis work at EPFL.
In particular, I am deeply grateful to Mats Granath, my supervisor and examiner,
for always responding quickly whenever I had questions or uncertainties and for
giving me clear guidance. He also trusted and supported my decisions during the
research and writing process. This support meant a great deal to me, especially
while I was working in Switzerland and handling both academic and administrative
matters far away from Chalmers.

I am also very grateful to my supervisor Sanidhya Kashyap at EPFL RS3Lab. His
support provided the experimental conditions that made this work on agent memory
possible, and his guidance during the research and paper writing process helped
shape the direction of the thesis. I would also like to thank EPFL and RS3Lab for
providing a stimulating research environment in which this project could be
carried out.

I owe special thanks to Yueyang Pan, a PhD student at RS3Lab, with whom I worked
closely on the agent memory project that became the foundation of this thesis.
Together, we pushed through the most intense weeks before the NeurIPS deadline,
supporting each other and completing the work under significant pressure. I will
not forget those three weeks of writing, experiments, debugging, and revisions. I
am equally thankful to Haodong Zheng, who worked with us on the paper and
contributed to bringing the project to completion.

I would also like to thank my girlfriend, Yang Li, for her constant care and
support. During the periods when I was anxious, overwhelmed, or extremely busy,
she was always there for me, listened to me, and helped me regain balance. Her
quiet support in the background has been one of the most important sources of
strength during this thesis.

Finally, I am deeply grateful to my family. Although we have been separated by a
great distance while I studied abroad, their unwavering support has always been
with me. Being able to study and work with peace of mind in another country would
not have been possible without them.

\vspace{1.5cm}
\hfill
\ThesisAuthor, Gothenburg, \ThesisMonthYear

\newpage				% Create empty back of side
\thispagestyle{empty}
\mbox{}
