\chapter{Supplementary Experimental Details}
\label{app:experimental-details}

This appendix records the experimental details that are compressed in
Chapter~\ref{chap:experiment-analysis}. The goal is to make the benchmark
construction, model configuration, training setup, and baseline adaptations
explicit enough for the results to be audited and reproduced.

\section{Training and Evaluation Split}
\label{app:training-evaluation-split}

The GRPO training data is constructed from three sources. LoCoMo and
LongMemEval provide public multi-session dialogue-memory questions with gold
answers and evidence-session annotations. The diagnostic \bench{} benchmark
contributes temporal-bridge records whose final answer evidence is not directly
addressable from the original query.

We start from 1,986 LoCoMo questions, 500 LongMemEval questions, and 500 raw
\bench{} candidates. For LoCoMo, category-5 adversarial questions are excluded.
For LongMemEval, single-session preference questions are excluded. For
\bench{}, the generation pipeline removes invalid attempts, duplicates, split
leakage, incomplete metadata, temporal inconsistencies, and non-unique answers.
This produces 1,536 LoCoMo questions, 469 LongMemEval questions, and 374
\bench{} records in the filtered candidate pool.

The final GRPO training set contains 500 examples. To avoid conversation-level
leakage, LoCoMo is partitioned by dialogue rather than by individual question:
six conversations form the LoCoMo training pool, and four conversations are
reserved for evaluation. LongMemEval is not used for GRPO training; all 469
filtered LongMemEval questions are used for zero-shot evaluation. For
\bench{}, 99 filtered records are sampled for training, and the remaining 275
records are used for evaluation. The evaluation partitions are not used for
checkpoint selection.

\begin{table}[H]
  \centering
  \caption{Training and evaluation split used for GRPO training and reporting.}
  \label{tab:training-evaluation-split}
  \scriptsize
  \resizebox{\textwidth}{!}{%
  \begin{tabular}{p{0.24\textwidth}rrrrp{0.24\textwidth}}
    \hline\hline
    Data stage & LoCoMo & LongMemEval & \bench{} & Total & Use \\
    \hline
    Original questions & 1,986 & 500 & 500 & 2,986 & Source pool \\
    Filtered questions & 1,536 & 469 & 374 & 2,379 & Candidate pool \\
    Training & 401 & 0 & 99 & 500 & GRPO training \\
    Evaluation & 655 & 469 & 275 & 1,399 & Reporting only \\
    \hline\hline
  \end{tabular}
  }
\end{table}

Each training example stores lightweight supervision and metadata: the user
query, gold answer, gold evidence sessions, source identifier, user or speaker
metadata, source bucket, difficulty label, and evaluation mode. It does not
store fixed retrieved memories or precomputed rollout traces. During GRPO, the
sample metadata is used to call the corresponding online memory-search
environment, retrieve sessions, synthesize events, produce verdicts, and score
the final answer. The policy is therefore trained on online memory-search
behavior rather than on a static retrieved context.

\section{Model and Baseline Configurations}
\label{app:model-combo}

Table~\ref{tab:model-combo} reports the model configuration used by each method
in the main result tables. Every method has a model that produces the final
answer. The distinction is whether the memory component returns evidence to a
separate downstream answerer, as \sys{}, Mem0, and A-Mem do, or whether the
same model directly produces the answer as part of the method, as in
full-context prompting and direct-answerer retrieval agents. GPT-5.1 is used as
the fixed judge for all methods and is not the answerer unless the row
explicitly names GPT-5.1 as the final-answer model.

\begin{table}[H]
  \centering
  \caption{Model combinations used in the main evaluation tables. The
  final-answer model is the model whose output is judged for answer accuracy.
  GPT-5.1 is the fixed judge for all methods, not the answerer for all methods.}
  \label{tab:model-combo}
  \scriptsize
  \resizebox{\textwidth}{!}{%
  \begin{tabular}{p{0.14\textwidth}p{0.17\textwidth}p{0.18\textwidth}p{0.25\textwidth}p{0.18\textwidth}}
    \hline\hline
    Group & Method & Output mode & Memory/search or policy model & Final-answer model \\
    \hline
    Full context & GPT-5.1 & Direct answer & Full history prompt & GPT-5.1 \\
    Full context & Qwen3.5-9B & Direct answer & Full history prompt & Qwen3.5-9B \\
    Write-time memory & Mem0 & Evidence provider & Qwen3.5-9B & Qwen3.5-9B \\
    Write-time memory & A-Mem & Evidence provider & Qwen3.5-9B & Qwen3.5-9B \\
    Search-time retrieval & MemR$^3$ & Direct answer & Qwen3.5-9B & Qwen3.5-9B \\
    Search-time retrieval & Search-R1 & Direct answer & Public checkpoint & Public checkpoint \\
    Temporal memory & Memory-T1-Qwen & Direct answer & Qwen3.5-9B & Qwen3.5-9B \\
    Ours & \sys{}-Vanilla & Evidence provider & Qwen3.5-9B & Qwen3.5-9B \\
    Ours & \sys{}-GRPO & Evidence provider & Qwen3.5-9B GRPO & Qwen3.5-9B \\
    \hline\hline
  \end{tabular}
  }
\end{table}

\section{Training Implementation Details}
\label{app:training-hyperparameters}

\textbf{Policy and rollout.}
The trainable policy is the Qwen3.5-9B Synthesizer, trained with
full-parameter GRPO using VERL/HybridFlow \cite{sheng2024hybridflow}. The
Planner, \sys{} retriever, Qwen3.5-9B answerer, GPT-5.1 judge, and reference
policy are frozen environment components. For each prompt, the actor samples
\(K=8\) trajectories with at most three retrieval rounds. At inference time,
\sys{} uses the same maximum of three retrieval rounds and retrieves the top-7
memories in each round. The rollout context allows 49,152 prompt tokens, 2,048
response tokens, and a 65,536-token SGLang model context
\cite{zheng2023sglang}.

\textbf{Optimization.}
We optimize with AdamW using learning rate \(2{\times}10^{-6}\), weight decay
0.01, betas \((0.9,0.999)\), and Adam epsilon \(10^{-8}\). The learning-rate
schedule is constant with a 3\% warmup, corresponding to 9 warmup steps over
300 global steps. Each GRPO update uses one prompt and eight rollout
trajectories, with PPO mini-batch size 1 and micro-batch size 1. We use PPO
clipping ratio 0.2, actor KL loss coefficient 0.001 with low-variance KL,
gradient clipping 1.0, bf16 training, and FSDP parameter and optimizer
offload. The final run resumes from global step 170 and continues to step 300,
saving checkpoints every 10 steps. Training is conducted on four NVIDIA GH200
96GB GPUs.

\textbf{Reward setting.}
The final run uses reward weights
\[
  (w_o,w_v,w_l)=(0.50,0.10,0.05),
  \qquad c=-0.20.
\]
Appendix~\ref{app:reward-design} defines the reward components, and
Appendix~\ref{app:reward-design-coefficients} analyzes the coefficient choice.

\section{Evaluation Protocol}
\label{app:evaluation-protocol}

The main metric is answer accuracy. In all experiments, GPT-5.1 is used as the
fixed LLM judge for automatic evaluation, and the judge prompt is kept
unchanged across compared methods. BLEU and token-level F1 are not used as
primary metrics because many long-term memory answers are semantically
equivalent despite surface variation, especially for dates, event descriptions,
and short natural-language answers.

LoCoMo is a two-speaker conversational memory benchmark. The evaluation follows
the conversation-memory setting and searches the relevant speaker histories as
required by each question. LongMemEval and \bench{} are treated as single-user
query settings; each query is evaluated independently over its associated
history or generated session set.

Token counts are output-side compactness metrics and are not identical to full
serving cost. For direct-answerer methods, tokens denote final answer-output
tokens. For evidence-provider methods, tokens denote the retrieved evidence or
memory context returned downstream. \sys{}-specific diagnostics additionally
include retrieval rounds and verdict accuracy. Retrieval rounds measure how
often the loop is used, while verdict accuracy measures whether the
Synthesizer's \texttt{sufficient}/\texttt{insufficient} decision agrees with
gold evidence coverage.

\section{Baseline Implementation Notes}
\label{app:baseline-implementation-notes}

This section records implementation adaptations and caveats not captured by
Table~\ref{tab:model-combo}. For stateful memory systems, users and benchmark
files are isolated to avoid cross-example leakage. Local Qwen baselines use
SGLang \cite{zheng2023sglang} with thinking disabled. Embedding-based
baselines use \texttt{text-embedding-3-large} with hidden dimension 3072 unless
otherwise noted. The intended evidence retrieval budget is top-\(k=7\), except
for method-specific budgets or implementation caveats below.

\textbf{Full context.}
Full context does not call a memory system. LoCoMo uses each conversation file
as one full-history prefix. LongMemEval uses each query-specific haystack, and
\bench{} provides all sessions in a sample as context. This is a strong
reference baseline but not a deployable memory system.

\textbf{Mem0.}
We run a local OSS Mem0 codebase \cite{chhikara2025mem0} with lightweight
server/client adaptations. Histories are ingested at the session level, and
Mem0 internally extracts fact-style memory notes; search therefore returns
Mem0 notes rather than raw transcripts. The local backend uses pgvector without
graph-store retrieval and adds synchronous ingestion, JSON repair and retry,
custom fact/update prompts, and top-\(k\) search fixes required by the local
evaluation harness.

\textbf{A-Mem.}
We evaluate A-Mem \cite{xu2025amem} with the released codebase. Each session
is inserted through one \texttt{add\_note()} call with a timestamped header, so
the stored \texttt{MemoryNote.content} is the full transcript rather than an
extracted atomic fact. The vector store is ChromaDB, and the wrapper keeps the
released add/search/evolution path intact. A-Mem evolution updates tags,
context, and links, but does not rewrite the stored transcript into compact
facts; its evidence token counts therefore reflect transcript-level evidence.

\textbf{MemR\(^3\).}
We evaluate MemR\(^3\) \cite{du2025memr3} with the released codebase and its
original retrieve/generate/reflect/answer workflow. Each benchmark item is
converted into MemR\(^3\)'s conversation-plus-QA format: the conversation is
flattened as timestamped speaker text, chunked, embedded, and queried by the
multi-step retrieval loop. The configuration uses chunk size 256, top-\(k=7\),
maximum 5 iterations, and no reranker. We do not train MemR\(^3\). Disabling
the reranker may understate its best retrieval performance, and its score
reflects retrieval plus answer generation rather than pure retrieval accuracy.

\textbf{Search-R1.}
We evaluate Search-R1 \cite{jin2025searchr1} with the public Hugging Face
checkpoint \texttt{PeterJinGo/SearchR1-nq\_hotpotqa\_train-qwen2.5-7b-em-ppo}.
The model is used as a direct answerer with the original
\texttt{<think>}/\texttt{<search>}/\texttt{<answer>} protocol. It can issue up
to three search calls, receives retrieved memory text in \texttt{<information>}
blocks, and the final text inside \texttt{<answer>} is used as the benchmark
response. The checkpoint was trained for open-domain search QA rather than
timestamped long-term conversational memory; the adapter does not fine-tune it
or extract atomic memory facts during ingestion.

\textbf{Memory-T1-Qwen.}
We do not run an official trained Memory-T1 checkpoint because public weights
were not available for the trained model at the time of the experiments. We
therefore report a Memory-T1-style baseline using Qwen3.5-9B rather than
official Memory-T1 checkpoint performance. Session-level SimpleBM25 retrieves
candidate session transcripts; the prompt asks Qwen to select memory IDs and
produce a JSON answer. The retrieval budget is top-\(k=7\), the prompt budget
is 15,500 tokens, temperature is 0, and malformed JSON outputs receive at most
two retries. Session-level BM25 can miss one endpoint of a temporal bridge, and
Qwen3.5-9B is not trained for Memory-T1-style long-term temporal dialogue
reasoning.

\chapter{Algorithm and Reward Details}
\label{app:algorithm-reward-details}

\section{Algorithm Details}
\input{include/backmatter/appendix_neurips/algorithm_details}

\chapter{Dataset Details}
\label{app:dataset-details}

\section{Dataset Construction}
\input{include/backmatter/appendix_neurips/dataset_details}

\chapter{Prompt Templates and Case Study}
\label{app:prompt-templates-and-case-study}

\section{Prompt Templates}
\input{include/backmatter/appendix_neurips/prompt_templates}
